---
title: "Case 2 - Experimental Design"
author: "Khoa Binh Tran"
date: "7/20/2020"
output: html_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE)
```
# Question

# Preparations
```{r loading_packages, message = FALSE}
library(dplyr)
library(ggplot2)
library(readr)
library(broom)
```
# Data
```{r}
lendingclub <- read_csv("datasets/case2_lendclub.csv")
lendingclub_ab <- read_csv("datasets/case2_lendclub_ab.csv")
```
#Analysis
##EDA
<!-- Examine the variables with glimpse() -->
```{r}
glimpse(lendingclub)
```
<!-- Find median loan_amnt and mean int_rate, annual_inc with summarize() -->
```{r}
lendingclub %>%
  summarize(median(loan_amnt),
            mean(int_rate),
            mean(annual_inc)
  )
```
<!-- Use ggplot2 to build a bar chart of purpose -->
```{r}
ggplot(lendingclub, aes(x = purpose)) +
    geom_bar() +
    coord_flip()
```
<!-- # Use recode() to create the new purpose_recode variable -->
```{r}
lendingclub$purpose_recode <- lendingclub$purpose %>% recode(
        "credit_card" = "debt_related",
          "debt_consolidation" = "debt_related",
          "medical" = "debt_related",
        "car" = "big_purchase",
          "major_purchase" = "big_purchase",
          "vacation" = "big_purchase",
        "moving" = "life_change",
          "small_business" = "life_change",
          "wedding" = "life_change",
        "house" = "home_related",
          "home_improvement" = "home_related",
          "purpose" = "purpose_recode")
```
##How does loan purpose affect amount funded?
<!-- Build a linear regression model, purpose_recode_model -->
```{r}
purpose_recode_model <- lm(funded_amnt ~ purpose_recode, data = lendingclub)
```
<!-- Examine results of purpose_recode_model -->
```{r}
summary(purpose_recode_model)
```
<!-- Get anova results and save as purpose_recode_anova -->
```{r}
purpose_recode_anova <- anova(purpose_recode_model)
```
<!-- Print purpose_recode_anova -->
```{r}
purpose_recode_anova
```
<!-- Examine class of purpose_recode_anova -->
```{r}
class(purpose_recode_anova)
```
Based on the very small p-value, purpose_recode_anova's results indicate that there is evidence to support the hypothesis that the mean loan amounts are different for at least one combination of purpose_recode's levels.

##Which loan purpose mean is different?
<!-- Use aov() to build purpose_aov -->
```{r}
purpose_aov <- aov(funded_amnt ~ purpose_recode, data = lendingclub)
```
<!-- Conduct Tukey's HSD test to create tukey_output -->
```{r}
tukey_output <- TukeyHSD(purpose_aov, "purpose_recode", conf.level = 0.95)
```
<!-- Tidy tukey_output to make sense of the results -->
```{r}
tidy(tukey_output)
```
Looking at the p-values for each comparison of the levels of purpose_recode, we can see that only a few of the mean differences are statistically significant, for example the differences in the means for the debt_related and big_purchase loan amounts. In this case, these tiny p-values are most likely to be due to large sample size, and further tests would be required to determine what's actually significant in the case of loans

##Multiple Factor Experiments
<!-- Use aov() to build purpose_emp_aov -->
```{r}
purpose_emp_aov <- aov(funded_amnt ~ purpose_recode + emp_length, data= lendingclub)
```
<!-- Print purpose_emp_aov to the console -->
```{r}
purpose_emp_aov
```
<!-- Call summary() to see the p-values -->
```{r}
summary(purpose_emp_aov)
```
##Model Validation
Examine what effect their Lending Club-assigned loan grade variable has on the interest rate, int_rate
###Pre-modeling EDA
<!-- Examine the summary of int_rate -->
```{r}
summary(lendingclub$int_rate)
```
<!-- Examine int_rate by grade -->
```{r}
lendingclub %>%
    group_by(grade) %>%
    summarize(mean = mean(int_rate), var = var(int_rate), median = median(int_rate))
```
<!-- Make a boxplot of int_rate by grade -->
```{r}
ggplot(lendingclub, aes(x = grade, y = int_rate)) +
    geom_boxplot()
```
<!-- Use aov() to create grade_aov plus call summary() to print results -->
```{r}
grade_aov <- aov(int_rate ~ grade, data = lendingclub)
summary(grade_aov)
```

You can see from the numeric summary and the boxplot that grade seems to heavily influence interest rate.

###Post-modeling validation plots + variance
Examining the Residuals vs. Fitted and Normal Q-Q plots.

<!-- For a 2x2 grid of plots: -->
```{r}
par(mfrow=c(2, 2))
```
<!-- Plot grade_aov -->
```{r}
plot(grade_aov)
```
<!-- Bartlett's test for homogeneity of variance -->
```{r}
bartlett.test(int_rate ~ grade, data = lendingclub)
```
The residuals on this model are okay, though the residuals on G have a much smaller range than any other level of grade (the dots are far less spread out.) The Q-Q plot, however, shows that the residuals are fairly normal. However, given the highly significant p-value from Bartlett's test, the assumption of homogeneity of variances is violated, which is one of the assumptions of an ANOVA model.

###Kruskal-Wallis rank sum test
The homogeneity of variance assumption of linear modeling was violated, we may want to try an alternative.
One non-parametric alternative to ANOVA is the Kruskal-Wallis rank sum test. The null hypothesis for this test would be that all of the int_rates have the same ranking by grade.

<!-- Conduct the Kruskal-Wallis rank sum test -->
```{r}
kruskal.test(int_rate ~ grade,
             data = lendingclub)
```
low p-value -> int_rate varies by grade

##A/B Testing
Does the color scheme of the header affect people's decision to apply for loans?

[A/B Test. Source: DataCamp](datasets/case2_ABtest.png)

###Sample size for A/B Test
Analyze A/B test results with t-test after collecting data. Calculate the sample size required.

Given the effect size d = 0.2, power 0.8, significance level 0.05 

<!-- Load the pwr package -->
```{r}
library(pwr)
```
<!-- Find the sample size -->
```{r}
pwr.t.test(n = NULL,                    
           d = 0.2,
           sig.level = 0.05,
           power = 0.8,
           alternative = "two.sided")
```
So, the sample size needed for the experiment is about 393 observations.

###A/B Test
When applicants were using the Lending Club website, they were randomly assigned to two groups, A or B, where A was shown a mint green website header and B was shown a light blue website header. Lending Club was interested to see if website header color choice influenced loan_amnt, the amount an applicant asked to borrow.
The A/B test was run until there were 500 applicants in each group. Each applicant has been labeled as group A or B. Data was collected in the lendingclub_ab dataset.

<!-- Plot the A/B test results -->
```{r}
ggplot(lendingclub_ab, aes(x = Group, y = loan_amnt)) +
    geom_boxplot()
```
<!-- Conduct a two-sided t-test -->
```{r}
t.test(loan_amnt ~ Group, data = lendingclub_ab)
```
It seems that there is no compelling evidence to support the hypothesis that there is a difference the two A/B test groups' mean loan_amnt.

Examine the effects of three explanatory variables on loan_amnt
<!-- Build lendingclub_multi -->
```{r}
lendingclub_multi <-lm(loan_amnt ~ Group + grade + verification_status, data = lendingclub_ab)
```
<!-- Examine lendingclub_multi results -->
```{r}
tidy(lendingclub_multi)
```