---
title: "Case Study 1 - Multiple Regression"
author: "Khoa Binh Tran"
date: "7/20/2020"
output: html_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE)
```
## Question
What are the factors that contribute to a price of the meal in a restaurant in NYC?

## Preparations
```{r loading_packages, message = FALSE}
library(dplyr)
library(ggplot2)
library(readr)
library(scatterplot3d)
```
## Data
```{r}
nyc <- read_csv("datasets/case1_nyc.csv")
```
## Analysis
### Exploratory Data Analysis
<!-- Exploring the data -->
```{r}
glimpse(nyc)
```
Examine the pairwise scatterplots
```{r}
nyc %>%
  select(-c("Case", "Restaurant")) %>%
  pairs()
```
###SLR Model
<!-- Price by Food plot -->
```{r}
ggplot(data = nyc, aes(x = Food, y = Price)) +
  geom_point()
```
<!-- Price by Food model -->
```{r}
mod1 <- lm(Price ~ Food, data = nyc)
summary(mod1)
```
Model 1 shows that for an additional unit in food quality, the price increases by approx. 2.94 dollars.
###The effect of location
Now we integrate a location variable (East) into the model. It is a categorical variable that specifies that whether the restaurant is located in the East(1) or not (0).
<!-- Summary statistics -->
```{r}
nyc %>%
  group_by(East) %>%
  summarize(mean(Price))
```
<!-- Parallel slopes model -->
```{r}
mod2 <- lm(Price ~ Food + East, data = nyc)
summary(mod2)
```
Each additional rating point of food quality is associated with a $2.88 increase in the expected price of meal, after controlling for location.
The premium for an Italian restaurant in NYC associated with being on the east side of 5th Avenue is $1.46, after controlling for the quality of the food.
###How about service?
Are people willing to pay more for better restaurant service?
<!-- Fit model -->
```{r}
mod3 <- lm(Price ~ Food + Service, data = nyc)
summary(mod3)
```
<!-- Draw 3D scatterplot -->
```{r}
p <- scatterplot3d(nyc$Price ~ nyc$Food + nyc$Service)
p$plane3d(mod3)
```
##Build a full model
Now we incorporate all three available variables in a model to explain for the variations of Price. (with exceptions to Case and Restaurant because they are id and name variables)
<!-- Price by Food and Service and Location -->
```{r}
mod4 <- lm(Price ~ Food + Service + East, data = nyc)
summary(mod4)
```
The premium for being on the East side of 5th Avenue is just less than a dollar, after controlling for the quality of food and service.

The impact of location is relatively small, since one additional rating point of either food or service would result in a higher expected price than moving a restaurant from the West side to the East side.

 <!-- draw 3D scatterplot -->
```{r}
library(plotly)
plot_ly(data = nyc, z = ~Price, x = ~Food, y = ~Service, opacity = 0.6) %>%
  add_markers(color = ~factor(East))
```
There is another variable that we havent considered yet: decor. Let's try fitting that into our model.
```{r}
mod5 <- lm(Price ~ Food + Service + Decor + East, data = nyc)
summary(mod5)
```
The effect of service turns negative, albeit statistically insignificant.

The dramatic change in the effect of service may be explained by the following reasons:

Since the quality of food, decor, and service were all strongly correlated, multicollinearity is the likely explanation.
Once we control for the quality of food, decor, and location, the additional information conveyed by service is negligible.